Finite element analysis ofmulti-dimensional and simplifiedmodels for beams and plates

### 2023-10-09 files/Overview.tex
Summary v1

The second part of the chapter considers work done in a textbook \cite{SF97}.

\subsubsection{Chapter 4}
\subsubsection{Chapter 5}
\subsubsection{Chapter 6}

>>>>

The second part of the chapter considers work done in a textbook \cite{SF97}. The specific work discussed, covers the convergence of the eigenvalues and eigenfunctions of a one-dimensional vibration problem when applying the Finite Element Method. The authors consider a general eigenvalue problem. Then using the Rayleigh-quotient, from the Rayleigh-Ritz method, as well as an approximation theorem from \cite{ORXX}, the main result is proven. The specific work done in this section is updating the notation of the textbook, as well as expanding some results so that the results are easier to understand.

\subsubsection{Chapter 4}
This chapter is a focus on the main model theory of this disseration, the Timoshenko beam theory. The first section is a discussion of modal analysis applied to the Timoshenko beam theory, and specifically a discussion of the article \cite{VV06}. In this article, the authors present a method to calculate the exact eigenvalues and eigenfunctions of a Timoshenko beam. This method considers the boundary conditions, which differs from most other research. Starting with a general eigenvalue problem for a Timoshenko beam model, the authors derive a general solution for this problem. The authors then explain the method by hands of an example by applying the method to a cantilever beam model. The next sections of this chapter also then looks at examples of applying the method, first to a cantilever beam model, and then to a free-free beam model. The focus of these applications are to show that the method works and that the numerical results can be obtained easily. Similar numerical results will be required in chapter 6. The next section is a short discussion on a suspended beam model, afterwards an interesting article \cite{SP06} is discussed. In this article, the authors investigate the validity of the Timoshenko beam theory, by comparing the eigenvalues (natural frequencies) for a emperically mesasured beam, to a Timoshenko beam and a three-dimensional beam using Finite Element Analysis. The specific model used is a free-free beam. The emperical results are obtained by vibrating a suspended steel beam and measuring the natural frequencies. These results are then compared to the theoretical results. Finally the chapter finished by looking at an interesting model of a cantilever beam with a tip-body with elastic interfaces. These results combine modal analysis and looks at the behavior of the model. This section serves as an example of how useful the Timoshenko beam theory can be in a complex model.

\subsubsection{Chapter 5}
In this chapter, the Finite Element Method is applied to cantilever two-dimensional,and three-dimensional cantilever beams and a cantilever Reissner-Mindlin plate. The aim of this section is to obtain a numerical formula to calculate the eigenvalues and eigenfunctions of the models. The results of these formulas will be used in chapter 6. The Finite Element Method need not be applied to the Timoshenko beam theory, as chapter 4 provides an alternative method. For all the models, the Finite Element Method is applied using bi-cubic or tri-cubic basis functions to improve the rate of convergence and reduce the processing required to obtain accurate results.

\subsubsection{Chapter 6}
In this chapter, the main comparisons of the models are conducted. The first section is a discussion of the article \cite{LVV09}. Each section in this chapter follows the same structure as this article. In this article, the authors investigate the validity of a cantilever Timoshenko beam models, by comparing it to a two-dimensional cantilever Timoshenko beam model. To make this comparison, the authors compare the eigenvalues and eigenfunctions of the models. But since the two-dimensional model is more complex, there are `non-beam type' eigenvalues. These are eigenvalue that are not in the Timoshenko beam theory, and does not relate to beam type problems. To be able to find the beam type eigenvalues that can be compared, the authors compare the mode shapes. This allows the eigenvalues to be matched up. The results of the article show that the Timoshenko beam model compares very well to the two-dimensional beam. The comparison improves when the application is for a long and slender beam. The results of the article are verified and more results are added to extend the article and also to be using the the following sections. But in real-world applications, a beam is a three-dimensional model. Therefore it would be better to use a three-dimensional model to investigate the validity of the Timoshenko beam theory. The authors of \cite{LVV09} mentions this, and suggests using the two-dimensional model as an intermediate step to avoid complexities. So in the second section of this chapter, the validity of the two-dimensional model is investigated using the three-dimensional model as a reference. The results show that the two-dimensional model also compares very well to the three-dimensional model. Similar to the one-dimensional case, the results improve for a long and slender beam, and also if the width of the three-dimensional beam is not larger than the height of the beam. When increasing the width past the size of the height, the two-dimensional model is less accurate. This shape does however bring into question the use of a beam model, and something like a plate model might be better suited. In the last section, the validiy of the Reissner-Mindlin plate model is investigated. Similar to how the Timoshenko beam theory improves on the Euler-Bernoulli beam theory, the Reissner-Mindlin plate theory improves on the classical plate theory. The numerical results do show that this Reissner-Mindlin plate model compares well to a three-dimensional plate model, also when the dimensions of the models are realistic for application.
